### 2016-06-07 guelph_talk/guelph_panther.Rnw
working on Guelph talk

\documentclass[english]{beamer}
\definecolor{links}{HTML}{2A1B81}
\hypersetup{colorlinks,linkcolor=,urlcolor=links}
\usepackage{natbib}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{multicol}
\usepackage{color}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{animate}
\usepackage{arydshln}
\usepackage[normalem]{ulem}
\newcommand{\colemph}[1]{{\color{red} \emph{#1}}}
%\bibliographystyle{shortreflist}
\bibliographystyle{notitle}

\usetheme{Frankfurt}
\usecolortheme{dove}
\setbeamercovered{transparent}
\setbeamercolor{description item}{fg=blue}
\newcommand{\citepx}[1]{{\small \citep{#1}}}
\newcommand{\citex}[1]{{\small \cite{#1}}}
% \renewcommand{\descriptionlabel}[1]{\hspace{\labelsep}\textbf{#1}}
\usepackage{babel}

% "Beamer infolines outer theme with miniframe bullets only for the current section"
% (http://tex.stackexchange.com/a/45152/3323)
%\useoutertheme[subsection=false]{miniframes}
\useoutertheme[subsection=false]{miniframes}
%% \setbeamertemplate{mini frame in other subsection}{}
%% \usepackage{etoolbox}
\makeatletter
\beamer@compressfalse
%% \patchcmd{\insertnavigation}{\hskip-1.875ex plus-1fill}{}{}{}
%% \patchcmd{\sectionentry}{\beamer@section@set@min@width}{}{}{}
%% \patchcmd{\sectionentry}{\hskip1.875ex plus 1fill}{}{}{}
%% \patchcmd{\sectionentry}{\hyperlink{Navigation#3}{{\usebeamertemplate{section in head/foot shaded}}}}{}{}{}
%% \patchcmd{\slideentry}{\beamer@ypos=#2\relax}{}{}{}
%% \patchcmd{\fakeslideentry}{\beamer@ypos=#2\relax}{}{}{}
\makeatother

\begin{document}

\newcommand{\rzero}{{\mathcal R}_0}
\newcommand{\rzeroi}[1]{{\mathcal R}_{0,#1}}

\makeatletter
\def\newblock{\beamer@newblock}
\makeatother

<<setup,echo=FALSE,message=FALSE>>=
library(knitr)
opts_chunk$set(echo=FALSE,error=FALSE,
               fig.align="center",out.width="0.8\\textwidth",
               fig.width=6,fig.height=4)
library(deSolve)
library(reshape)
library(lattice)
library(latticeExtra)
library(emdbook)
library(nlme)
library(bbmle)
library(tikzDevice)
## http://yihui.name/en/2011/04/produce-authentic-math-formulas-in-r-graphics/
options(tikzMetricPackages = c( "\\usepackage[utf8]{inputenc}",
                "\\usepackage[T1]{fontenc}", "\\usetikzlibrary{calc}",
                "\\usepackage{amssymb}"))

##options(tikzMetricPackages = c("\\usepackage[utf8]{inputenc}",
##    "\\usepackage[T1]{fontenc}", "\\usetikzlibrary{calc}",
##    "\\usepackage{amssymb}"))
library(ggplot2)
theme_set(theme_bw())
library("ggplot2"); theme_set(theme_classic())
library("MASS")
## use Dark2 rather than Set1 because colour #6 of Set1 is yellow (ugh/too light)
scale_colour_discrete <- function(...,palette="Dark2") scale_colour_brewer(...,palette=palette)
scale_fill_discrete <- function(...,palette="Dark2") scale_fill_brewer(...,palette=palette)
library("gridExtra")
library("plyr")  ## for ldply()
library("dplyr") ## for full_join()
library("reshape2")  ## for melt()
library("GGally")
library("ggstance")
library(scales) ## for squish()
library(gridExtra) ## for grid.arrange()
library(mgcv)
library(plyr)   ## for mutate()
library(psyphy) ## for restricted-link models
library(reshape2)
library(splines)  ## for vir asymptote profile
library(deSolve)
library(Hmisc)
library(gsubfn)
@

\title[Animal movement]{Model complexity and model choice for animal movement models}
\author[Ben Bolker]{Ben~Bolker, McMaster University \\
  {\tiny Departments of Mathematics \& Statistics and Biology}}
\institute{Guelph Biomathematics \& Biostatistics Symposium}
\date{9 June 2016}
% \pgfdeclareimage[height=0.5cm]{uflogo}{letterhdwm}
% \logo{\pgfuseimage{uflogo}}
\AtBeginSubsection[]{
  \frame<beamer>{ 
     \frametitle{Outline}   
     \tableofcontents[currentsection,currentsubsection] 
   }
 }

\begin{frame}
\titlepage
\end{frame}
% \beamerdefaultoverlayspecification{<+->}

\begin{frame}
\frametitle{Outline}
\tableofcontents{}
\end{frame}

\begin{frame}
\frametitle{Acknowledgements}
\begin{description}
\item[People] Michael Li, Madelon van de Kerk, Dave Onorato
\item[Support] NSERC Discovery grant
\end{description}

\end{frame}

\begin{frame}
\frametitle{To do}
\begin{itemize}
\item pix!
\item general REFS: Turchin, Morales et al, Langrock ... ?
\end{itemize}
\end{frame}

\section{Animal movement}
\subsection{}

\begin{frame}
\frametitle{Animal movement: data}

\begin{itemize}
\item observations:  \\
e.g. mass mark-recapture, longitudinal density,
direct observation, telemetry (VHF, GPS)
\item most methods provide a sequence of times and locations for each individual
\item summaries:
\begin{itemize}
\item home range \\
(convex hull, kernel density estimate, etc.)
\item root-mean-squared displacement
\item step length and turning angle
\end{itemize}
\item covariates: \\
e.g. habitat map, individual characteristics (sex, age, weight \ldots)
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Animal movement: questions}
\begin{itemize}
\item simple description
\item how do animals' movements change  
as a function of their (internal or external) environment?  \\
what does that tell us about their biology?
\item how might animals' distributions, etc. change when conditions 
(density, habitat, ...) change?
\end{itemize}
\end{frame}

\section[Panthers]{Florida panthers}
\subsection{}

\begin{frame}
\frametitle{Biological/conservation issues}

\begin{columns}
  \begin{column}{0.5\textwidth}
    \begin{itemize}
    \item Florida panther: \emph{Felix ...}
    \item severely reduced habitat, endangered
    \item small, isolated population
    \item currently recovering
    \end{itemize}
  \end{column}
\end{columns}

\end{frame}

\begin{frame}
\frametitle{Panther movement questions}

\begin{columns}
  \begin{column}{0.5\textwidth}
    \begin{itemize}
    \item abc
    \end{itemize}
  \end{column}
\end{columns}

\end{frame}

\begin{frame}
\frametitle{Panther movement data}

\begin{columns}
  \begin{column}{0.5\textwidth}
    \begin{itemize}
    \item panthers tracked, captured
    \item GPS collars
    \item 18 males (13 male, 5 female, 1-15 years old)
    \item 3200 panther days, hourly/bihourly; 49000 locations
    \item ?? per panther
    \end{itemize}
  \end{column}
  \begin{column}{0.5\textwidth}
    \includegraphics[trim={10cm 0 10cm 0},clip,width=\textwidth]{pix/panther_collar.jpg}
  \end{column}
\end{columns}
\end{frame}

\begin{frame}
\frametitle{example movement tracks}
FIXME: picture here!
(animation?? map background??)
\end{frame}

\section[HMM]{Hidden Markov models}
\subsection{}

\begin{frame}
\frametitle{Hidden Markov models}

\begin{itemize}
\item finite mixture model with temporal dependence
\item discrete time steps
\item discrete latent state; \emph{transition matrix}
\item observations from \emph{emission distributions} \\
(continuous or discrete, univariate or multivariate)
\end{itemize}

\includegraphics[width=\textwidth]{pix/hmm.jpg}
\end{frame}

\begin{frame}
\frametitle{Hidden Markov models (cont.)}

\begin{itemize}
\item \emph{forward-backward algorithm} for estimating parameters
\item \emph{Viterbi algorithm} for estimating most probable state sequences
\item {\tt depmixS4} package \citepx{visser2010depmixs4}
(also {\tt moveHMM} \citepx{michelot_movehmm_2016}, {\tt move.HMM} \citepx{augustine_hmm})
\end{itemize}

\end{frame}

\section[Basic analysis]{Basic analysis (vdK et al.)}
\subsection{}

\begin{frame}
\frametitle{State distributions}

\includegraphics[trim={0 6.5cm 0 0}, clip, width=\textwidth]{pix/jae_fig2.png}

\end{frame}

\begin{frame}
\frametitle{Parameter estimates}

\includegraphics[width=\textwidth]{pix/jae_fig3.png}

\end{frame}

\begin{frame}
\frametitle{Transition parameters}

picture/table here of transition parameters (network diagram??)
\end{frame}

\begin{frame}
\frametitle{Diurnal variation}

\includegraphics[trim={0 8cm 0 0}, clip, width=\textwidth]{pix/jae_fig4.png}

\end{frame}

\begin{frame}
\frametitle{what can we conclude so far?}

\textbf{good news}

\begin{itemize}
\item basic biology: males move faster, farther
\item three states are identifiable, sensible
\item ...
\end{itemize}

\pause
\textbf{bad news}
\begin{itemize}
\item diurnal variation in Viterbi results - but it's not in the model!
\item estimates of model complexity are too high
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Model complexity}

\includegraphics[height=\textheight]{pix/jae_fig1.png}

\end{frame}

\section[Diurnal model]{Incorporating diurnal variation}
\subsection{}

\begin{frame}
\frametitle{Broadening the model}

Attempting to fix these problems:
\begin{itemize}
\item extend the model to allow covariates
\item specifically, allow for diurnal variation
\begin{itemize}
\item simplify model (log-Normal step length only)
\item \emph{fixed} state-specific emissions parameters\\
(step length mean and std dev)
\item time-varying transition parameters
\item also try \emph{finite mixture models} \\
(independent occupancy)
\end{itemize}
\item how much does this help?
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Temporal models}

figure showing alternative temporal models
\end{frame}

\begin{frame}
\frametitle{Comparison}

Figure 1 from paper
\end{frame}

\begin{frame}
\frametitle{Goodness of fit}

Figure 2 from paper
\end{frame}

\begin{frame}
\frametitle{Temporal patterns}
Figures 3 (and 4?) from paper
\end{frame}

\begin{frame}
\frametitle{Model complexity}
Figure showing BIC plots for all cats tried
\end{frame}

\begin{frame}
\frametitle{Model complexity}
Simulation results
\end{frame}

\section{Broader issues/outlook}
\subsection{}

\begin{frame}
\frametitle{Big data and small models}

\begin{itemize}
\item simple model families +  \\
  model misspecification $\to$ \\
  overparameterization

\end{itemize}
\end{frame}

\begin{frame}
\frametitle{an aside on AIC vs BIC}

\end{frame}

\begin{frame}
\frametitle{Tools needed}

\begin{itemize}
\item cross-validation \citepx{wenger_assessing_2012}
\item diagnostic plots
\item score tests?
\end{itemize}

\end{frame}

\begin{frame}
\begin{multicols}{2}
\frametitle{References}

% \tiny
\fontsize{3pt}{5pt}\selectfont
\bibliography{../paper}
\end{multicols}
\end{frame}

\end{document}