### 2016-08-02 bmc_template/figures.Rnw
need figure capations and fix up cat 14 and 15

\documentclass{article}

\begin{document}

<<pkgs,message=FALSE,warning=FALSE,echo=FALSE>>=
library("ggplot2"); theme_set(theme_bw())
scale_colour_discrete <- function(...,palette="Set1")
    scale_colour_brewer(...,palette=palette)
scale_fill_discrete <- function(...,palette="Set1")
    scale_fill_brewer(...,palette=palette)
zmargin <- theme(panel.margin=grid::unit(0,"lines"))
library("knitr")
opts_chunk$set(fig.width=6,fig.height=4)
## fix complaints about "incompatible color definition" from xcolor ...
## http://tex.stackexchange.com/questions/148188/knitr-xcolor-incompatible-color-definition
knit_hooks$set(document = function(x) {sub('\\usepackage[]{color}', '\\usepackage{xcolor}', x, fixed = TRUE)})
library("depmixS4")
library("plyr")
library("dplyr")
library("reshape2")
library("gridExtra")
library("RColorBrewer")
if (!require("ggalt") || length(find("geom_encircle"))==0) {
#     cat("installing ggalt from BB repo ...\n")
    try(detach("package:ggalt"),silent=TRUE)
    devtools::install_github("hrbrmstr/ggalt")
    library("ggalt")
}
@

<<sim_results, warning=FALSE,echo=FALSE,message=FALSE,fig.cap="Simulation test for BIC-optimal approach",fig.width=6,fig.height=4>>=
ICdat <- readRDS("../sim_results/ICdat.rds")
aa <- readRDS("../sim_results/sim.101.RDS")

ICdat <- mutate(ICdat,
                model=factor(model,
                         levels=c("HMM","HMMsin"),
                         labels=c("homogeneous","heterogeneous"))) %>% filter(IC=="BIC")
simplot <- ggplot(ICdat,aes(x=nstates,y=n,color=model,
                            linetype=IC,shape=IC))+
                        geom_point(size=3) +
                        geom_line() +
                        scale_colour_brewer(palette="Set1")+
                        labs(y="counts",x="# latent states")
print(simplot)

@

\clearpage

<<BICred_plot,warning=FALSE,echo=FALSE,message=FALSE,fig.width=6,fig.height=4>>=
allmodsumdf2 <- readRDS("../summary_stats/cat.1.df.RDS")$multimoddf
df2 <- transform(allmodsumdf2, #change allmodsumdf to allmodsumdf2
            distrib=ifelse(grepl("W",model),"Weibull","log-Normal"),
            turn_angle=ifelse(grepl("VM",as.character(model)),"yes","no"),
            homog=ifelse(grepl("TH",type),"time-geneous HMM","homogeneous HMM"))
(ggplot(df2, aes(x=nstates,y=deltaBIC,colour=distrib,lty=turn_angle))
    + geom_line()
    + facet_wrap(~ homog)
    + scale_colour_brewer(name="step length\ndistribution",palette="Set1")
    ## scale_size_continuous( name="# parameters") +
    #    geom_point(aes(size=parameters))+
    + scale_linetype(name="von Mises\nturning angle?")
    + labs(y=expression(Delta*"BIC"),
         x="Number of States")
    + geom_hline(yintercept=0,colour="gray")
    + zmargin)
@

\clearpage

<<adj_BIC_comparisons, warning=FALSE, echo=FALSE, message=FALSE,,fig.width=8>>=
#L <- load("cat1sumdat.RData")

linetypev <- c(rep(2,2),rep(1,5))
sumdat <- transform(readRDS("../summary_stats/cat.1.df.RDS")$HMMsumdf,
                     mod = ifelse(type=="HMM + TH","HMM + TH","FMMs and HMM")
)
linetypev <- c(rep(2,2),rep(1,5))
(adj_BIC_plot <- ggplot(sumdat, aes(x=parameters
                                  , y=deltaBIC
                                  , colour=model
                                  , linetype=model
                                  , group=model))
   + facet_wrap(~mod,ncol = 2,scales = "free_x")
   + scale_size_continuous(name="# of states")
   + scale_x_continuous(breaks=seq(0,300,by=30))
   + geom_point(aes(shape=factor(nstates),alpha=nstates))
   + scale_linetype_manual(values=linetypev, name="Model")
   + scale_colour_brewer(palette="Dark2", name="Model")
   + scale_shape_discrete(name="Number of states")
   + scale_alpha_continuous(name="Number of states")
   + labs(y=expression(paste(Delta, "BIC"))
          , x="Number of parameters") 
   + geom_line()
   + zmargin)
@

\clearpage

<<avg_step_length_by_time, message=FALSE, echo=FALSE, warning=FALSE,fig.width=8>>=

avgdf <- readRDS("../summary_stats/cat.1.df.RDS")$avgplotdf
obsdf <- (readRDS("../summary_stats/cat.1.RDS") 
  %>% dplyr::select(c(LogDist,Time)) 
  %>% filter(!is.na(LogDist)) 
  %>% group_by(Time) 
  %>% summarise_each(funs(mean))
  %>% transmute(time = Time, variable = "obs", value=LogDist)
)
vit <- readRDS("../summary_stats/cat.1.vit.RDS")
vitdf <- (vit %>% dplyr::select(c(viterbi,time))
  %>% group_by(time)
  %>% summarise_each(funs(mean))
  %>% mutate(model="homogeneous (n=3)")
)

vitdf2 <- vitdf %>% transmute(time=time,variable="viterbi",value=viterbi)

mnames <- c("fmm5","fmmsin4","hmm6","hmmblock4","hmmhourly3",
            "hmmquad5","hmmsin5","obs","viterbi")
mnames2 <- c("fmm-homogeneous  (n=5)"
             , "fmm-sin (n=4)"
             , "homogeneous (n=6)"
             , "block (n=4)"
             , "hourly (n=3)"
             , "quadratic (n=5)"
             , "sin (n=5)"
             , "observed"
             , "viterbi homogenous (n=3)")

combdf <- (rbind(avgdf,obsdf,vitdf2)
           %>% transmute(time=time,model=variable,steplen=value)
           %>% filter(model %in% mnames)
           %>% mutate(model=factor(model,levels=mnames,labels=mnames2))
)
linetypev <- c(rep(c(2,1),c(2,6))
                ,0
               )
colvec <- c(c(brewer.pal(7,"Dark2"),"black")
             ,"black"
            )
lwdvec <- c(rep(c(1,4),c(7,1))
             ,4
            )
alphavec <- c(rep(c(1,0.2),c(7,1))
               ,1
              )

shapevec <- c(rep(32,8),118)

(ggplot(combdf,aes(x=time,y=steplen,colour=model,
                   lwd=model,
                   alpha=model,linetype=model, shape=model))
  + geom_point()
  + geom_line()
  + xlab("Time of Day") 
  + scale_shape_manual(values=shapevec, name="Model")
  + ylab(expression('Mean step length'~(log[10]*m)))
  + scale_colour_manual(values=colvec, name="Model")
  + scale_alpha_manual(values=alphavec, name="Model")
  + scale_size_manual(values=lwdvec, name="Model")
  + scale_linetype_manual(values=linetypev, name="Model")
 ) 
@

\clearpage

<<acf_plot, message=FALSE, echo=FALSE, warning=FALSE>>=
acffun <- function(sim){
  a <- acf(sim,na.action=na.pass,plot=FALSE)
  df <- data.frame(ACF=c(a$acf),lag=a$lag)
  return(df)
}
vitacf <- (vit %>% dplyr::select(-c(time,obs))%>%ldply(.,acffun) %>% mutate(model=.id)
)

acfdf <- (readRDS("../summary_stats/cat.1.df.RDS")$acfdf 
          %>% mutate(model = .id)
)

combacfdf <- (rbind(acfdf,vitacf)
           %>% transmute(lag=lag,model=model,ACF=ACF)
           %>% filter(model %in% mnames)
           %>% mutate(model=factor(model,levels=mnames,labels=mnames2))
)

(ggplot(combacfdf,aes(x=lag,y=ACF,colour=model,
                  lwd=model,alpha=model,linetype=model,shape=model))
  + geom_point()
  + geom_line()
  + xlab("Lag") 
  + ylab("Autocorrelation")
  + scale_shape_manual(values=shapevec, name="Model")
  + scale_colour_manual(values=colvec, name="Model")
  + scale_alpha_manual(values=alphavec, name="Model")
  + scale_size_manual(values=lwdvec, name="Model")
  + scale_linetype_manual(values=linetypev, name="Model")
)

## vitacf messes up legend ...
@

\clearpage

<<r msdlist>>=
get_cat_pars2 <- function(catid){
  templist <- readRDS(paste("../summary_stats/cat.",catid,".df.RDS",sep=""))
  moddf <- templist$msdlist

  df <- data.frame(catID = catid
          , log_mean = c(unlist(moddf[1]),unlist(moddf[3]))
          , log_sd = c(unlist(moddf[2]),unlist(moddf[4]))
          , Model = c(rep("HMM",length(unlist(moddf[1]))),rep("HMM+TH(sin)",length(unlist(moddf[3]))))
  )
  return(df)
}

all_cat_pars2 <- ldply(lme4:::namedList(1,22,14,15),get_cat_pars2)

g0 <- ggplot(all_cat_pars2,aes(log_mean,log_sd,color=Model))+
     geom_point()+geom_line() + facet_wrap(~catID,labeller=label_both)+zmargin
 (g0c <- g0 +
     geom_encircle(data=subset(all_cat_pars2,
                               (catID=="15" & log_mean<1.5) |
                               (catID!="15" & log_mean<1.8)),
                   colour="black",linetype=2,expand=0.05))

@

\clearpage

\section{sup}

<<BICred_plot2,warning=FALSE,echo=FALSE,message=FALSE,fig.width=6,fig.height=4>>=

allmodsumdf2 <- readRDS("../summary_stats/cat.2.df.RDS")$multimoddf
df2 <- transform(allmodsumdf2, #change allmodsumdf to allmodsumdf2
            distrib=ifelse(grepl("W",model),"Weibull","log-Normal"),
            turn_angle=ifelse(grepl("VM",as.character(model)),"yes","no"),
            homog=ifelse(grepl("TH",type),"time-geneous HMM","homogeneous HMM"))
(ggplot(df2, aes(x=nstates,y=deltaBIC,colour=distrib,lty=turn_angle))
    + geom_line()
    + facet_wrap(~ homog)
    + scale_colour_brewer(name="step length\ndistribution",palette="Set1")
    ## scale_size_continuous( name="# parameters") +
    #    geom_point(aes(size=parameters))+
    + scale_linetype(name="von Mises\nturning angle?")
    + labs(y=expression(Delta*"BIC"),
         x="Number of States")
    + geom_hline(yintercept=0,colour="gray")
    + zmargin)
@

\clearpage

<<adj_BIC_comparisons2, warning=FALSE, echo=FALSE, message=FALSE,,fig.width=8>>=
#L <- load("cat1sumdat.RData")

linetypev <- c(rep(2,2),rep(1,5))
sumdat <- transform(readRDS("../summary_stats/cat.2.df.RDS")$HMMsumdf,
                     mod = ifelse(type=="HMM + TH","HMM + TH","FMMs and HMM")
)
linetypev <- c(rep(2,2),rep(1,5))
(adj_BIC_plot <- ggplot(sumdat, aes(x=parameters
                                  , y=deltaBIC
                                  , colour=model
                                  , linetype=model
                                  , group=model))
   + facet_wrap(~mod,ncol = 2,scales = "free_x")
   + scale_size_continuous(name="# of states")
   + scale_x_continuous(breaks=seq(0,300,by=30))
   + geom_point(aes(shape=factor(nstates),alpha=nstates))
   + scale_linetype_manual(values=linetypev, name="Model")
   + scale_colour_brewer(palette="Dark2", name="Model")
   + scale_shape_discrete(name="Number of states")
   + scale_alpha_continuous(name="Number of states")
   + labs(y=expression(paste(Delta, "BIC"))
          , x="Number of parameters") 
   + geom_line()
   + zmargin)
@

\clearpage

<<avg_step_length_by_time2, message=FALSE, echo=FALSE, warning=FALSE,fig.width=8>>=

avgdf <- readRDS("../summary_stats/cat.2.df.RDS")$avgplotdf
obsdf <- (readRDS("../summary_stats/cat.2.RDS") 
  %>% dplyr::select(c(LogDist,Time)) 
  %>% filter(!is.na(LogDist)) 
  %>% group_by(Time) 
  %>% summarise_each(funs(mean))
  %>% transmute(time = Time, variable = "obs", value=LogDist)
)
vit <- readRDS("../summary_stats/cat.2.vit.RDS")
vitdf <- (vit %>% dplyr::select(c(viterbi,time))
  %>% group_by(time)
  %>% summarise_each(funs(mean))
  %>% mutate(model="homogeneous (n=3)")
)

vitdf2 <- vitdf %>% transmute(time=time,variable="viterbi",value=viterbi)

mnames <- c("fmm4","fmmsin4","hmm5","hmmblock4","hmmhourly3",
            "hmmquad4","hmmsin4","obs","viterbi")
mnames2 <- c("fmm-homogeneous  (n=4)"
             , "fmm-sin (n=4)"
             , "homogeneous (n=5)"
             , "block (n=4)"
             , "hourly (n=3)"
             , "quadratic (n=4)"
             , "sin (n=4)"
             , "observed"
             , "viterbi homogenous (n=3)")

combdf <- (rbind(avgdf,obsdf,vitdf2)
           %>% transmute(time=time,model=variable,steplen=value)
           %>% filter(model %in% mnames)
           %>% mutate(model=factor(model,levels=mnames,labels=mnames2))
)
linetypev <- c(rep(c(2,1),c(2,6))
                ,0
               )
colvec <- c(c(brewer.pal(7,"Dark2"),"black")
             ,"black"
            )
lwdvec <- c(rep(c(1,4),c(7,1))
             ,4
            )
alphavec <- c(rep(c(1,0.2),c(7,1))
               ,1
              )

shapevec <- c(rep(32,8),118)

(ggplot(combdf,aes(x=time,y=steplen,colour=model,
                   lwd=model,
                   alpha=model,linetype=model, shape=model))
  + geom_point()
  + geom_line()
  + xlab("Time of Day") 
  + scale_shape_manual(values=shapevec, name="Model")
  + ylab(expression('Mean step length'~(log[10]*m)))
  + scale_colour_manual(values=colvec, name="Model")
  + scale_alpha_manual(values=alphavec, name="Model")
  + scale_size_manual(values=lwdvec, name="Model")
  + scale_linetype_manual(values=linetypev, name="Model")
 ) 
@

\clearpage

<<acf_plot2, message=FALSE, echo=FALSE, warning=FALSE>>=
acffun <- function(sim){
  a <- acf(sim,na.action=na.pass,plot=FALSE)
  df <- data.frame(ACF=c(a$acf),lag=a$lag)
  return(df)
}
vitacf <- (vit %>% dplyr::select(-c(time,obs))%>%ldply(.,acffun) %>% mutate(model=.id)
)

acfdf <- (readRDS("../summary_stats/cat.2.df.RDS")$acfdf 
          %>% mutate(model = .id)
)

combacfdf <- (rbind(acfdf,vitacf)
           %>% transmute(lag=lag,model=model,ACF=ACF)
           %>% filter(model %in% mnames)
           %>% mutate(model=factor(model,levels=mnames,labels=mnames2))
)

(ggplot(combacfdf,aes(x=lag,y=ACF,colour=model,
                  lwd=model,alpha=model,linetype=model,shape=model))
  + geom_point()
  + geom_line()
  + xlab("Lag") 
  + ylab("Autocorrelation")
  + scale_shape_manual(values=shapevec, name="Model")
  + scale_colour_manual(values=colvec, name="Model")
  + scale_alpha_manual(values=alphavec, name="Model")
  + scale_size_manual(values=lwdvec, name="Model")
  + scale_linetype_manual(values=linetypev, name="Model")
)

## vitacf messes up legend ...
@

\clearpage
%cat14

<<BICred_plot14,warning=FALSE,echo=FALSE,message=FALSE,fig.width=6,fig.height=4>>=

allmodsumdf2 <- readRDS("../summary_stats/cat.14.df2.RDS")$multimoddf
df2 <- transform(allmodsumdf2, #change allmodsumdf to allmodsumdf2
            distrib=ifelse(grepl("W",model),"Weibull","log-Normal"),
            turn_angle=ifelse(grepl("VM",as.character(model)),"yes","no"),
            homog=ifelse(grepl("TH",type),"time-geneous HMM","homogeneous HMM"))
(ggplot(df2, aes(x=nstates,y=deltaBIC,colour=distrib,lty=turn_angle))
    + geom_line()
    + facet_wrap(~ homog)
    + scale_colour_brewer(name="step length\ndistribution",palette="Set1")
    ## scale_size_continuous( name="# parameters") +
    #    geom_point(aes(size=parameters))+
    + scale_linetype(name="von Mises\nturning angle?")
    + labs(y=expression(Delta*"BIC"),
         x="Number of States")
    + geom_hline(yintercept=0,colour="gray")
    + zmargin)
@

\clearpage

<<adj_BIC_comparisons14, warning=FALSE, echo=FALSE, message=FALSE,,fig.width=8>>=
#L <- load("cat1sumdat.RData")

linetypev <- c(rep(2,2),rep(1,5))
sumdat <- transform(readRDS("../summary_stats/cat.14.df2.RDS")$HMMsumdf,
                     mod = ifelse(type=="HMM + TH","HMM + TH","FMMs and HMM")
)
linetypev <- c(rep(2,2),rep(1,5))
(adj_BIC_plot <- ggplot(sumdat, aes(x=parameters
                                  , y=deltaBIC
                                  , colour=model
                                  , linetype=model
                                  , group=model))
   + facet_wrap(~mod,ncol = 2,scales = "free_x")
   + scale_size_continuous(name="# of states")
   + scale_x_continuous(breaks=seq(0,300,by=30))
   + geom_point(aes(shape=factor(nstates),alpha=nstates))
   + scale_linetype_manual(values=linetypev, name="Model")
   + scale_colour_brewer(palette="Dark2", name="Model")
   + scale_shape_discrete(name="Number of states")
   + scale_alpha_continuous(name="Number of states")
   + labs(y=expression(paste(Delta, "BIC"))
          , x="Number of parameters") 
   + geom_line()
   + zmargin)
@

\clearpage

<<avg_step_length_by_time14, message=FALSE, echo=FALSE, warning=FALSE,fig.width=8>>=

avgdf <- readRDS("../summary_stats/cat.14.df2.RDS")$avgplotdf
obsdf <- (readRDS("../summary_stats/cat.14.RDS") 
  %>% dplyr::select(c(LogDist,Time)) 
  %>% filter(!is.na(LogDist)) 
  %>% group_by(Time) 
  %>% summarise_each(funs(mean))
  %>% transmute(time = Time, variable = "obs", value=LogDist)
)
vit <- readRDS("../summary_stats/cat.14.vit.RDS")
vitdf <- (vit %>% dplyr::select(c(viterbi,time))
  %>% group_by(time)
  %>% summarise_each(funs(mean))
  %>% mutate(model="homogeneous (n=3)")
)

vitdf2 <- vitdf %>% transmute(time=time,variable="viterbi",value=viterbi)

mnames <- c("fmm4","fmmsin4","hmm5","hmmblock4","hmmhourly3",
            "hmmquad4","hmmsin4","obs","viterbi")
mnames2 <- c("fmm-homogeneous  (n=4)"
             , "fmm-sin (n=4)"
             , "homogeneous (n=5)"
             , "block (n=4)"
             , "hourly (n=3)"
             , "quadratic (n=4)"
             , "sin (n=4)"
             , "observed"
             , "viterbi homogenous (n=3)")

combdf <- (rbind(avgdf,obsdf,vitdf2)
           %>% transmute(time=time,model=variable,steplen=value)
           %>% filter(model %in% mnames)
           %>% mutate(model=factor(model,levels=mnames,labels=mnames2))
)
linetypev <- c(rep(c(2,1),c(2,6))
                ,0
               )
colvec <- c(c(brewer.pal(7,"Dark2"),"black")
             ,"black"
            )
lwdvec <- c(rep(c(1,4),c(7,1))
             ,4
            )
alphavec <- c(rep(c(1,0.2),c(7,1))
               ,1
              )

shapevec <- c(rep(32,8),118)

(ggplot(combdf,aes(x=time,y=steplen,colour=model,
                   lwd=model,
                   alpha=model,linetype=model, shape=model))
  + geom_point()
  + geom_line()
  + xlab("Time of Day") 
  + scale_shape_manual(values=shapevec, name="Model")
  + ylab(expression('Mean step length'~(log[10]*m)))
  + scale_colour_manual(values=colvec, name="Model")
  + scale_alpha_manual(values=alphavec, name="Model")
  + scale_size_manual(values=lwdvec, name="Model")
  + scale_linetype_manual(values=linetypev, name="Model")
 ) 
@

\clearpage

<<acf_plot14, message=FALSE, echo=FALSE, warning=FALSE>>=
acffun <- function(sim){
  a <- acf(sim,na.action=na.pass,plot=FALSE)
  df <- data.frame(ACF=c(a$acf),lag=a$lag)
  return(df)
}
vitacf <- (vit %>% dplyr::select(-c(time,obs))%>%ldply(.,acffun) %>% mutate(model=.id)
)

acfdf <- (readRDS("../summary_stats/cat.14.df.RDS")$acfdf 
          %>% mutate(model = .id)
)

combacfdf <- (rbind(acfdf,vitacf)
           %>% transmute(lag=lag,model=model,ACF=ACF)
           %>% filter(model %in% mnames)
           %>% mutate(model=factor(model,levels=mnames,labels=mnames2))
)

(ggplot(combacfdf,aes(x=lag,y=ACF,colour=model,
                  lwd=model,alpha=model,linetype=model,shape=model))
  + geom_point()
  + geom_line()
  + xlab("Lag") 
  + ylab("Autocorrelation")
  + scale_shape_manual(values=shapevec, name="Model")
  + scale_colour_manual(values=colvec, name="Model")
  + scale_alpha_manual(values=alphavec, name="Model")
  + scale_size_manual(values=lwdvec, name="Model")
  + scale_linetype_manual(values=linetypev, name="Model")
)

## vitacf messes up legend ...
@

\end{document}